---
format: html
editor: visual
---

```{r, eval = F}
# This code chunk contains code to install all of the dependencies
# necessary to compile and run your report, using if-statements to
# reduce install time for unnecessary code.
# It should be set to eval = F by default, so you aren't installing
# software on someone's computer without their consent.

# This works for packages that are on CRAN
if (!"dplyr" %in% installed.packages()) {
  install.packages("dplyr")
}
if (!"remotes" %in% installed.packages()) {
  install.packages("remotes")
}

# This is how to do it for a package that's only on github
if (!"emo" %in% installed.packages()) {
  remotes::install_github("hadley/emo")
}  

```



```{r}
BabyNames <- read.csv("Popular_Baby_Names.csv")
#colnames(BabyNames)
#str(BabyNames)
#unique_Ethic <- unique(BabyNames$Ethnicity); unique_Ethic

sum(is.na(BabyNames)) # check missing values; no missing values

library(dplyr)
# distinct this dataset; remove the repeated rows based on X element 
#D_BabyNames <- BabyNames %>% distinct() 

#updated_data <- D_BabyNames %>% group_by(Ethnicity) %>% summarise()
# count frequency of each ethnic group
#updated_data1 <- D_BabyNames %>% group_by(Ethnicity) %>% count()

BabyNames$Child.s.First.Name <- tolower(BabyNames$Child.s.First.Name)

#BabyNames$Child.s.First.Name <- tolower(BabyNames$Child.s.First.Name)
# Using unique function to get the same dataset as D_BabyNames
babynames<- unique(BabyNames) # remove the repeated rows; 
#babynames$Child.s.First.Name <- tolower(babynames$Child.s.First.Name)

# combine the ethnic groups b/c they have the same meaning
modified_data <- babynames %>% mutate(Ethnicity = recode(Ethnicity, "BLACK NON HISP" = "BLACK NON HISPANIC" ,
                                                      "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                                                      "WHITE NON HISP" = "WHITE NON HISPANIC" )) 

# check the result; there are total 4 levels in the ethnic groups
#M_data1 <- modified_data %>% group_by(Ethnicity) %>% count()
#M_data2 <- modified_data %>% group_by(Ethnicity, Gender) %>% count()
```

## Introduction

### Description of the Data set

This dataset `BabyNames` consist of the popular baby names by sex and ethnic group from 2011 to 2019, and it was collected through civil birth registration. There is other information about how the data.citynewyork.us collected the data. But we think they may collect the data from online registration system. Each row represents **a** **popular baby** first name . There are 6 columns, `Year of Birth`, it is quantitative and indicates the year of the baby birth; `Year of Birth` includes year from 2011 to 2019; `Gender`, it is categorical variable (Male and Female); `Ethnicity` is a categorical variable and it indicates the mother's ethnicity; `Child.s.First.Name` is a categorical variable, it represents baby's first name; `Count` is quantitative and the value represents the frequency of names were collected through civil birth registration; and `Rank` represents the frequency of baby names in descending order in each year.

<<<<<<< HEAD
### Cleaning the Data

In the data cleaning phase, we imported the dataset and discovered that it does not contain any missing values. Additionally, we identified and removed duplicate rows from the initial dataset. There are some ethic groups have the similar names such as `ASIAN AND PACIFIC ISLANDER` and `ASIAN AND PACI`, `BLACK NON HISPANIC` and `BLACK NON HISP` and `WHITE NON HISPANIC` and `WHITE NON HISP`. They are the same ethic groups but with different names, so we combine them before we perform analysis regarding our interest, and finally we will use the `modified_data` to do further data analysis. `Black non-Hispanic(BNH)` refers to a person having origins in any of the black racial groups of Africa (except those of Hispanic origin); `White non-Hispanic(WNH)` means a person having origins in any of the original peoples of Europe, the Middle East, or North Africa but not in Hispanic. `Hispanic(H)` refers to a person of Cuban, Mexican, Puerto Rican, South or Central American, or other Spanish culture or origin, regardless of race.; and `Aisan and Pacific Islander(API)` means a person having origins in any of the original peoples of the Far East, Southeast Asia, or the Indian Subcontinent, including, for example, Cambodia, China, India, Japan, Korea, Malaysia, Pakistan, the Philippine Islands, Thailand, and Vietnam, also includes a person having origins in any of the original peoples of Hawaii, Guam, Samoa, or other Pacific Islands. Finally, there are total 2 quantitative variables, `Year of Birth` and `Count`; 4 categorical variables, `Gender`, `Ethnicity`, `Child.s.First.Name` and `Rank`. 


[Ethnic Information1](https://portal.ct.gov/-/media/DOT/documents/ddbe/RaceandEthnicIdentificationDefinitions2015pdf.pdf) 
[Ethnic Information2](https://research.schev.edu/info/Glossary/Black-or-African-American-Black-non-Hispanic#:~:text=Term%3A%20Black%20or%20African%20American,except%20those%20of%20Hispanic%20origin)
[Ethnic Information23](https://www.census.gov/topics/population/race/about.html)

=======

Firstly, we display the top ten popular first names in each gender and ethnic groups and we are going to use a bar chart to display it. People can easily read these two graphs. The top 10 popular baby's first names were displayed on the horizontal axis, the occurrence of each baby's first names was displayed on the vertical line. From the bar graphs regarding different ethnic groups, there are 4 bars charts for 4 different ethnic groups. It is easy to read the top 10 baby's first name from each bar group but it may not easy to read a more common results at once. We also plot a bar graph displaying top 10 baby's first name based on a joint data set.

We want to investigate whether the names that have consistently remained popular over several years or decades. In this question, we need to use inner_join function to see if we can get the common names over time. This analysis needs much time than previous questions. (note: 2011 to 2019, 9 years, summarize 9 dataframes with popular names between genders/ethnic groups) Display names in genders/ethnic groups each year in interactive graph, then display to merged data.


Besides, we will display a time plot to show the average length of first names recorded in the dataset from 2011 to 2019. Analyze any trends here such as the increasing or decreasing trend. Why it changes as time moves on. Short names are popular and this is a trend today. (note: analyze the average of length of first names between genders from 2011 to 2019) To calculate the length of each names, apply what we learned in the class and use dyply package to calculate the average length. Then display the time plot which the horizontal axis is time, the vertical axis is the average length of first names.


Draft of Samitha - How does the ranking (rank) of a name change over time? - A possible questions we might be able to answer is, are there any noticeable spikes in the popularity of names related to popular culture, such as movie characters or celebrities?

Feedback from professor: "It might be interesting to look at different spellings of similar names (e.g. Keily, Kylee, Kylie)" no need to identify all, some are good too.

## Methods

Use `str()` function to display the structure of the data, identify the categorical variables and quantitative variables. Use `sum()` to see any missing values in the data. There is no missing values and we also see some repeated rows, so we use `distinct()` to get unique rows in the raw data.

### Describe any data cleaning and rearranging you needed to do to get your dataset into a workable form. Make sure to cite any packages which were important in your data cleaning process in this section. For instance, if you used dplyr, then it would be appropriate to say something like

> we used the group-apply-combine paradigm with the `dplyr` functions `group_by` and `summarize` [@dplyr-package] to generate a dataset for each day of the observation period from the 15-minute interval observations in the raw data set.

## Topic of Exploration
The first question is what the top 10 popular first names are in each gender and ethnic groups. We use the bar chart to visualize this result. 

```{r filter dataset by M and F}
#| echo: false

library(dplyr)
# select top 10 female's names
vf_data <- modified_data %>% 
  filter(Gender == "FEMALE") %>% 
  group_by(Child.s.First.Name, Gender) %>% 
  summarise(freq = sum (Count),  .groups = "drop") %>% 
  arrange(desc(freq)) %>% head(10)

# select top 10 male's names
vm_data <- modified_data %>% 
  filter(Gender == "MALE") %>% 
  group_by(Child.s.First.Name, Gender) %>%
  summarise(freq = sum(Count), .groups = "drop") %>% 
  arrange(desc(freq)) %>% 
  head(10)
```


#####Figure 1.1: bar graph of top 10 popular baby's first name between girls and boys

```{r}
#| label: fig-genders1
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows the top 10 first names for males and females."
#| echo: false

library(ggplot2)

vf_data$Child.s.First.Name <-factor(vf_data$Child.s.First.Name, levels = vf_data$Child.s.First.Name) 

ggplot(data=vf_data, aes(x= Child.s.First.Name, y=freq)) +
  geom_bar(stat="identity", fill = 'hotpink') +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(title = "Top 10 Popular Female Babies' Names", 
       subtitle = "from 2011 to 2019", 
       x = "Child's First Name", 
       y = "Count") + 
    theme_minimal()

vm_data$Child.s.First.Name <-factor(vm_data$Child.s.First.Name, levels = vm_data$Child.s.First.Name) 

ggplot(data=vm_data, aes(x= Child.s.First.Name, y=freq)) +
  geom_bar(stat="identity", fill = 'steelblue2') +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(title = "Top 10 Popular Male Babies' Names", 
       subtitle = "from 2011 to 2019", 
       x = "Child's First Name", 
       y = "Count") + 
    theme_minimal()
```

#####Figure 1.2 bar graph of top 10 names in different ethnic groups for male and female

```{r filter dataset by ethnic groups}
#| echo: false

# top 10 names in ethnic group of API between Males and Females
em1_data <- modified_data %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER" & Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_API = sum(Count), .groups = "drop") %>% arrange(desc(freqM_API)) %>% head(n=10)

ef1_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_API = sum(Count)) %>% arrange(desc(freqF_API)) %>% slice_head(n=10)


# top 10 names in ethnic group of BNH between Males and Females
em2_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_BNH = sum(Count)) %>% arrange(desc(freqM_BNH)) %>% slice_head(n=10)

ef2_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_BNH = sum(Count)) %>% arrange(desc(freqF_BNH)) %>% slice_head(n=10)

# top 10 names in ethnic group of H between Males and Females
em3_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_H = sum(Count)) %>% arrange(desc(freqM_H)) %>% slice_head(n=10)

ef3_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_H = sum(Count)) %>% arrange(desc(freqF_H)) %>% slice_head(n=10)

# top 10 names in ethnic group of WNH between Males and Females
em4_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_WNH = sum(Count)) %>% arrange(desc(freqM_WNH)) %>% slice_head(n=10)

ef4_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_WNH = sum(Count)) %>% arrange(desc(freqF_WNH)) %>% slice_head(n=10)

```

#####Figure 1.2.1 bar graph of top 10 Asian and Pacific Islander boys names
```{r}
#| label: fig-genders2
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Asian and Pacific Islander boys names"
#| echo: false


# display in ordered by count
em1_data$Child.s.First.Name <-factor(em1_data$Child.s.First.Name, levels = em1_data$Child.s.First.Name) 

ggplot(data = em1_data, aes(x = Child.s.First.Name, y = freqM_API)) + 
  geom_bar(stat = "identity", fill = 'lightblue') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 Asian and Pacific Islander First Names For Boys",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.2 bar graph of top 10 Asian and Pacific Islander girls names
```{r}
#| label: fig-genders3
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Asian and Pacific Islander girls names"
#| echo: false


# display in ordered by count
ef1_data$Child.s.First.Name <-factor(ef1_data$Child.s.First.Name, levels = ef1_data$Child.s.First.Name) 

ggplot(data = ef1_data, aes(x = Child.s.First.Name, y = freqF_API)) + 
  geom_bar(stat = "identity", fill = 'lightpink') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10  Asian and Pacific Islander First Names For Girls",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```


#####Figure 1.2.3 bar graph of top 10 Black non Hispanic boys names
```{r}
#| label: fig-genders4
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Black non Hispanic boys names"
#| echo: false


# display in ordered by count
em2_data$Child.s.First.Name <-factor(em2_data$Child.s.First.Name, levels = em2_data$Child.s.First.Name) 

ggplot(data = em2_data, aes(x = Child.s.First.Name, y = freqM_BNH)) + 
  geom_bar(stat = "identity", fill = 'lightblue') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 Black non Hispanic First Names For Boys",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.4 bar graph of top 10 Black non Hispanic girls names
```{r}
#| label: fig-genders5
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Black non Hispanic girls names"
#| echo: false

# display in ordered by count
ef2_data$Child.s.First.Name <-factor(ef2_data$Child.s.First.Name, levels = ef2_data$Child.s.First.Name) 

ggplot(data = ef2_data, aes(x = Child.s.First.Name, y = freqF_BNH)) + 
  geom_bar(stat = "identity", fill = 'lightpink') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 Black non Hispanic First Names For Girls",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.5 bar graph of top 10 Hispanic boys names
```{r}
#| label: fig-genders6
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Hispanic boys names"
#| echo: false


# display in ordered by count
em3_data$Child.s.First.Name <-factor(em3_data$Child.s.First.Name, levels = em3_data$Child.s.First.Name) 

ggplot(data = em3_data, aes(x = Child.s.First.Name, y = freqM_H)) + 
  geom_bar(stat = "identity", fill = 'lightblue') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 Hispanic First Names For Boys",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.6 bar graph of top 10 Hispanic girls names
```{r}
#| label: fig-genders7
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular Hispanic girls names"
#| echo: false


# display in ordered by count
ef3_data$Child.s.First.Name <-factor(ef3_data$Child.s.First.Name, levels = ef3_data$Child.s.First.Name) 

ggplot(data = ef3_data, aes(x = Child.s.First.Name, y = freqF_H)) + 
  geom_bar(stat = "identity", fill = 'lightpink') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 Hispanic First Names For Girls",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.7 bar graph of top 10 White Non Hispanic boys names
```{r}
#| label: fig-genders8
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 popular White Non Hispanic boys names"
#| echo: false

# display in ordered by count
em4_data$Child.s.First.Name <-factor(em4_data$Child.s.First.Name, levels = em4_data$Child.s.First.Name) 

ggplot(data = em4_data, aes(x = Child.s.First.Name, y = freqM_WNH)) + 
  geom_bar(stat = "identity", fill = 'lightblue') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 White Non Hispanic First Names For Boys",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```

#####Figure 1.2.8 bar graph of top 10 White Non Hispanic girls names
```{r}
#| label: fig-genders9
#| fig-width: 8
#| fig-height: 4 # Change the figure dimensions using fig-width/height
#| out-width: 80% # This changes the size of the figure as rendered in the text. 
#| fig-cap: "This figure shows top 10 White Non Hispanic girls names"
#| echo: false


# display in ordered by count
ef4_data$Child.s.First.Name <-factor(ef4_data$Child.s.First.Name, levels = ef4_data$Child.s.First.Name) 

ggplot(data = ef4_data, aes(x = Child.s.First.Name, y = freqF_WNH)) + 
  geom_bar(stat = "identity", fill = 'lightpink') +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Top 10 White Non Hispanic First Names For Girls",
       x = "Name",
       y = "Count") + 
      theme_minimal()
```


From each bar chart, it is easy to read the top 10 names between males and females which are collected through civil birth registration on the website. Besides, the bar charts also display the top 10 babies names for different ethnic groups. These bar charts are easily to read by gender and ethnic groups individually. However, how to visualize the common names among different ethnic groups? We use the `inner_join` function to merge the two datasets for two ethnic groups to develop whether there are some common results. 

Here, we merge the datasets consist of White non-Hispanic (WNH) Males and Black non-Hispanic (BNH) Males, and the White non-Hispanic (WNH) Females and Black non-Hispanic (BNH) Females, dnd then we use the bar charts to plot the `common` results.
```{r merge data sets for all males and females in different 4 ethnic groups 10 to 20}
#| echo: false

library(dplyr)

# develop merge of males between 2 ethnic groups
inner_M1 <- inner_join(em1_data, em3_data, by = "Child.s.First.Name")
inner_M2 <- inner_join(em2_data, em4_data, by = "Child.s.First.Name") 
inner_M2 # no common result; why? different culture

# expand dataset with top 20 names
em22_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_BNH = sum(Count)) %>% arrange(desc(freqM_BNH)) %>% slice_head(n=20)

em44_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_WNH = sum(Count)) %>% arrange(desc(freqM_WNH)) %>% slice_head(n=20)


inner_M22 <- inner_join(em22_data, em44_data, by = "Child.s.First.Name")
inner_M22 # luckily, we get common results but the culture plays an important role in names

####---------------------------------------------------------------------------------------------####

# develop merge of female between 2 ethnic groups
inner_F1 <- inner_join(ef1_data, ef3_data, by = "Child.s.First.Name")
inner_F2 <- inner_join(ef2_data, ef4_data, by = "Child.s.First.Name")
inner_F2 # only one common names

ef22_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_BNH = sum(Count)) %>% arrange(desc(freqF_BNH)) %>% slice_head(n=20)

ef44_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_WNH = sum(Count)) %>% arrange(desc(freqF_WNH)) %>% slice_head(n=20)

inner_F22 <- inner_join(ef22_data, ef44_data, by = "Child.s.First.Name")
inner_F22 # Again, cultures difference 

####-----------------------------------------------------------------------####
```


```{r merge data sets for all males and females in different 4 ethnic groups 50}
#| echo: false


# expand dataset with top 50 names
em50_BNH_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_BNH = sum(Count)) %>% arrange(desc(freqM_BNH)) %>% slice_head(n=50)

em50_WNP_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_WNH = sum(Count)) %>% arrange(desc(freqM_WNH)) %>% slice_head(n=50)


inner_M50 <- inner_join(em50_BNH_data, em50_WNP_data, by = "Child.s.First.Name")
inner_M50 # more common names but not too much

####---------------------------------------------------------------------------------------------####


ef50_BNH_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_BNH = sum(Count)) %>% arrange(desc(freqF_BNH)) %>% slice_head(n=50)

ef50_WNH_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_WNH = sum(Count)) %>% arrange(desc(freqF_WNH)) %>% slice_head(n=50)

inner_F50 <- inner_join(ef50_BNH_data, ef50_WNH_data, by = "Child.s.First.Name")
inner_F50 # Again, cultures difference 

####-----------------------------------------------------------------------####
```

After we merge the dataset of WNH and BNH for males and females for top 10 names, there is no common result for males and only one common name (Olivia) for female. The main reason is the different culture between BNH and WNH. We extend the top 10 to top 20, 3 boys names are used and 4 girls names are used for both ethnic groups. Even though we can see more common results but the frequency of the names are very different. We plot a side-by-side bar chart to show the frequency between WNH and BNH. What we really want to tell the audience is different names are used in different ethnic groups, and one reason is because of different cultures. 


#####Figure 1.3.1  side by side bar graph of common names in API and H for male
```{r}
#| label: fig-ethnic1
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows common names for Asian Pacific Islander and Hispanic (MALES)"
#| echo: false

library(reshape2)
melt_M1 <- melt(inner_M1, id.vars = 'Child.s.First.Name')
melt_M1$variable <- as.character(melt_M1$variable )
melt_M1$variable <- gsub('freqM_API', 'Asian Pacific Islander - Male', melt_M1$variable)
melt_M1$variable <- gsub('freqM_H', 'Hispanic - Male', melt_M1$variable)

ggplot(data = melt_M1, aes(x = Child.s.First.Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Common Males Frist Name Between Asian Pacific Islander and Hispanic",
        subtitle = "within top 10", 
       x = "Name",
       y = "Frequency",
       fill = "Ethnic Group")+
  theme_minimal()

```

#####Figure 1.3.2  side by side bar graph of common names in WNH and BNH for male (top 20)
```{r}
#| label: fig-ethnic2
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows common names for Black non-Hispanic and White non-Hispanic (MALES) top 20"
#| echo: false


melt_M22 <- melt(inner_M22, id.vars = 'Child.s.First.Name')
melt_M22$variable <- as.character(melt_M22$variable )
melt_M22$variable <- gsub('freqM_BNH', 'Black non-Hispanic - Male', melt_M22$variable)
melt_M22$variable <- gsub('freqM_WNH', 'White non-Hispanic - Male', melt_M22$variable)

ggplot(data = melt_M22, aes(x = Child.s.First.Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Common Males Frist Name Between Black non-Hispanic and White non-Hispanic",
        subtitle = "within top 20", 
       x = "Name",
       y = "Frequency",
       fill = "Ethnic Group")+
  theme_minimal()

```


#####Figure 1.3.3  side by side bar graph of common names in API and H for female
```{r}
#| label: fig-ethnic3
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows common names for Asian Pacific Islander and Hispanic (FEMALES)"
#| echo: false


inner_F1 <- melt(inner_F1, id.vars = 'Child.s.First.Name')
inner_F1$variable <- as.character(inner_F1$variable )
inner_F1$variable <- gsub('freqF_API', 'Asian Pacific Islander - Female', inner_F1$variable)
inner_F1$variable <- gsub('freqF_H', 'Hispanic - Female', inner_F1$variable)

ggplot(data = inner_F1, aes(x = Child.s.First.Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Common Female Frist Name Between Asian Pacific Islander and Hispanic",
        subtitle = "within top 10", 
       x = "Name",
       y = "Frequency",
       fill = "Ethnic Group") +
  theme_minimal()
```

#####Figure 1.3.2  side by side bar graph of common names in WNH and BNH for female 
```{r}
#| label: fig-ethnic4
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows common names for Black non-Hispanic and White non-Hispanic (FEMALES)" 
#| echo: false


melt_F2 <- melt(inner_F2, id.vars = 'Child.s.First.Name')
melt_F2$variable <- as.character(melt_F2$variable )
melt_F2$variable <- gsub('freqF_BNH', 'Black non-Hispanic - Female', melt_F2$variable)
melt_F2$variable <- gsub('freqF_WNH', 'White non-Hispanic - Female', melt_F2$variable)

ggplot(data = melt_F2, aes(x = Child.s.First.Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Common Males Frist Name Between Black non-Hispanic and White non-Hispanic",
        subtitle = "within top 10", 
       x = "Name",
       y = "Frequency",
       fill = "Ethnic Group")+
  theme_minimal()

```

#####Figure 1.3.2  side by side bar graph of common names in WNH and BNH for female (top 20)
```{r}
#| label: fig-ethnic5
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows common names for Black non-Hispanic and White non-Hispanic (FEMALES) top 20"
#| echo: false


melt_F22 <- melt(inner_F22, id.vars = 'Child.s.First.Name')
melt_F22$variable <- as.character(melt_F22$variable )
melt_F22$variable <- gsub('freqF_BNH', 'Black non-Hispanic - Female', melt_F22$variable)
melt_F22$variable <- gsub('freqF_WNH', 'White non-Hispanic - Female', melt_F22$variable)

ggplot(data = melt_F22, aes(x = Child.s.First.Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Common Males Frist Name Between Black non-Hispanic and White non-Hispanic",
        subtitle = "within top 20", 
       x = "Name",
       y = "Frequency",
       fill = "Ethnic Group")+
  theme_minimal()

```


`word clouds (interactive graph; ref: 12; plot the UNISEX name)`
```{r Develop Unisex Name: merge males and females within the same ethnic groups}
#| echo: false

## Develop Unisex Name: merge males and females within the same ethnic groups 

em1_100_data <- modified_data %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER" & Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_API = sum(Count), .groups = "drop") %>% arrange(desc(freqM_API)) %>% head(n=1000)

ef1_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_API = sum(Count)) %>% arrange(desc(freqF_API)) %>% slice_head(n=1000)

inner_com1 <- inner_join(em1_100_data, ef1_100_data, by = "Child.s.First.Name")
# found 1 unisex name


em2_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_BNH = sum(Count)) %>% arrange(desc(freqM_BNH)) %>% slice_head(n=1000)

ef2_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_BNH = sum(Count)) %>% arrange(desc(freqF_BNH)) %>% slice_head(n=1000)

inner_com2 <- inner_join(em2_100_data, ef2_100_data, by = "Child.s.First.Name")
# no result

### selected 1000 to see whether there is unisex name; check M_data2 to each # of each gender or ethnic group
em3_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_H = sum(Count)) %>% arrange(desc(freqM_H)) %>% slice_head(n=1000)

ef3_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_H = sum(Count)) %>% arrange(desc(freqF_H)) %>% slice_head(n=1000)

inner_com3 <- inner_join(em3_100_data, ef3_100_data, by = "Child.s.First.Name")


em4_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_WNH = sum(Count)) %>% arrange(desc(freqM_WNH)) %>% slice_head(n=1000)

ef4_100_data <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_WNH = sum(Count)) %>% arrange(desc(freqF_WNH)) %>% slice_head(n=1000)

inner_com4 <- inner_join(em4_100_data, ef4_100_data, by = "Child.s.First.Name")

```

#### Time Plot of Average Length of First Names by Gender
```{r calculate average length of letters for names}
#| echo: false

library(stringr)
# calculate the length of letter of each name
length_data <- modified_data %>% mutate(N.Length = str_count(Child.s.First.Name, "[A-Za-z]")) %>% mutate(T.length = Count * N.Length) 
# calculate the average number of letters in each year for males and females
avglength_data <- length_data  %>% group_by(Year.of.Birth, Gender) %>% 
  summarise(Avg.Letters = mean(N.Length, na.rm = TRUE), .groups = "drop") %>% 
  arrange(Year.of.Birth, Gender)
# calculate the average number of letters in each year for each ethnic groups
avglength_data_ethnic <- length_data  %>% group_by(Year.of.Birth, Ethnicity) %>% 
  summarise(Avg.Letters = mean(N.Length, na.rm = TRUE), .groups = "drop") %>% 
  arrange(Year.of.Birth, Ethnicity)

```


#####Time Plot 1.1 Average Length of First Names For Male and Female
```{r}
#| label: fig-AverageLength1
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows average length of babies' first name for female and male from 2011 to 2019"
#| echo: false


ggplot(data = avglength_data, aes(x = Year.of.Birth, y = Avg.Letters, color = Gender, group = Gender)) + 
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2011, 2019, by=1), limits = c(2011, 2019)) +
  labs(title = "Average Number of Letters in Babies' Name by Gender", 
       subtitle = "from 2011 to 2019", 
       x = "Year", 
       y = "Average Number of Letters") + 
    theme_minimal()
```

#####Time Plot 1.2 Average Length of First Names For Different Ethnic Groups
```{r}
#| label: fig-AverageLength2
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "This figure shows average length of babies' first name for different ethnic groups from 2011 to 2019"
#| echo: false


ggplot(data = avglength_data_ethnic, aes(x = Year.of.Birth, y = Avg.Letters, color = Ethnicity, group = Ethnicity)) + 
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2011, 2019, by=1), limits = c(2011, 2019)) +
  labs(title = "Average Number of Letters in Babies' Name by Ethnicity", 
       subtitle = "from 2011 to 2019", 
       x = "Year", 
       y = "Average Number of Letters") + 
    theme_minimal()
```

From the time plot of average length of letters for babies first names, we can see, overall, the average length of names decreases over the time. The main reason is trend. Today, more people prefer short name because it make easier for people to remember and short name can also make it unique. 

### Need to modify.... Cannot remember where I found this reference

```{r}
#| echo: false

#babynames %>%
 # filter(name %in% c("William", "Will", "Bill") & sex== "M") %>%
#  ggplot(aes(year, n, colour=name)) + geom_line() + 
 # theme_economist() +
#  xlab("Year") + 
 # ylab("Number of People") +
#  ggtitle("Changes in Popularity of the Name William")


```

#### THIS IS ORIGINAL CODING FROM qmd. FILE

The `echo: false` option disables the printing of code (only output is displayed). If you generate a figure, it should have a caption. Here's a demonstration of how to do that:


Then, you can reference @fig-iris in the text and the appropriate cross-reference will be generated.

You can find additional information about formatting figures generated from code in the [quarto documentation](https://quarto.org/docs/authoring/figures.html#computations).

## Additional Exploration topic



It might be interesting to look at different spellings of similar names (e.g. Keily, Kylee, Kylie), if you get a chance and want to challenge yourselves. I'm not sure if you can get all sets of similar spellings automatically (maybe with some sort of phonetic guide or data read in from elsewhere, but that would be a bit complicated) but you could just look at a few common variations that pop up from your EDA.

`Add another topic here... as many as you desire, really. Make sure to include a transition between the two sections that connects the two with some sort of logical train of thought.`

## Conclusion





## Tips

(delete this section from your report!)

Almost anything you might want to know about how to format output in quarto can be found [here](https://quarto.org/docs/authoring/markdown-basics.html). Feel free to email/come to office hours to figure out how to do XYZ - part of the goal of making you write this report is that I want you to know how to write e.g. a journal paper in Quarto as well, so now's the time to experiment.

If you want to know what the wordcount of your report is, you can run the following command in your terminal:

```         
pandoc --lua-filter wordcount.lua report.qmd
```

Notice that I have not pushed `_output/report.html` or the `_output/report_files/` folder to github - this is intentional. I have actually set `_output` to not show up in git, to encourage you all to NOT push the rendered files to github and to instead work from the markdown files directly.

You may find it cleaner to create a figure subdirectory and store any figures that aren't created by R/Python in that folder. I encourage you to organize this repository in a sensible way.
