---
title: "S_work"
format: html
editor: visual
---

Libraries used

```{r, echo=FALSE}
library(tidyverse)
library(reshape2)
library(phonics)

#install.packages("wordcloud")
library(wordcloud)
#install.packages("RColorBrewer")
library(RColorBrewer)
#install.packages("wordcloud2")
library(wordcloud2)
#install.packages("tm")
library(tm)
install.packages("ggwordcloud")
library(ggwordcloud)

library ('plyr') 
```

Load the data

```{r}
BabyNames <- read.csv("Popular_Baby_Names.csv")
```

Data Cleaning

```{r}

#check missing values; no missing values
sum(is.na(BabyNames)) 

#combine the ethnic groups b/c they have the same meaning
modified_data<- BabyNames %>% mutate(Ethnicity = recode(Ethnicity, 
"BLACK NON HISP" = "BLACK NON HISPANIC" ,
"ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
"WHITE NON HISP" = "WHITE NON HISPANIC" )) 



#remove the repeated rows
babynames<- unique(modified_data) 

modified_data <- babynames


```

Summary of the data set

```{r}
colnames(babynames)
str(babynames)

babynames %>% group_by(Ethnicity) %>% summarise(count = n())
babynames %>% group_by(Ethnicity, Gender) %>% summarise(count = n())

babynames$Child.s.First.Name <- tolower(babynames$Child.s.First.Name)

```

Draft of Samitha - How does the ranking (rank) of a name change over time? - A possible questions we might be able to answer is, are there any noticeable spikes in the popularity of names related to popular culture, such as movie characters or celebrities?

Feedback from professor: "It might be interesting to look at different spellings of similar names (e.g. Keily, Kylee, Kylie)" no need to identify all, some are good too.

Create a Word Cloud (I will do it later after working with the rank variable)

```         
```

```         
```

Analysis on the rank variable

```{r}
top10_names <- modified_data %>% filter(Rank %in% c(1:10))
```

```{r, warning=FALSE}

#Most popular name for the males in the each ethinc group

df1 <- top10_names %>% filter(Ethnicity == "HISPANIC" & Gender == "MALE") %>% arrange(Rank) %>% head(9)


ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular names for Male Hispanics ", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")




df1 <- top10_names %>% filter(Ethnicity == "WHITE NON HISPANIC"  & Gender == "MALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular names for Male White Non-Hispanics ", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")


        
df1 <- top10_names %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER"  & Gender == "MALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular names for  Male Asian Pacific Islanders", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")   



df1 <- top10_names %>% filter(Ethnicity == "BLACK NON HISPANIC"  & Gender == "MALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular names forMale Black Non-Hispanics", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")   
  




```

```{r, warning=FALSE}
#Most popular name for the females in the each ethinc group

df1 <- top10_names %>% filter(Ethnicity == "HISPANIC" & Gender == "FEMALE") %>% arrange(Rank) %>% head(9)


ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular name for FEMALE Hispanics ", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")




df1 <- top10_names %>% filter(Ethnicity == "WHITE NON HISPANIC"  & Gender == "FEMALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular name for FEMALE White Non-Hispanics ", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")


        
df1 <- top10_names %>% filter(Ethnicity == "ASIAN AND PACIFIC ISLANDER"  & Gender == "FEMALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular name for  FEMALE Asian Pacific Islanders", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")   



df1 <- top10_names %>% filter(Ethnicity == "BLACK NON HISPANIC"  & Gender == "FEMALE") %>% arrange(Rank) %>% head(9)

ggplot(df1, aes(x = Year.of.Birth, y = Count )) +
  geom_line(color = "red", linetype = "solid", size = 1.25) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.25, color = "blue", size = 4) +
  geom_text(aes(label = c(Count)), nudge_x = 0.25, nudge_y = 5, color = "black", size = 3.5) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  labs(title = "Most popular name for FEMALE Black Non-Hispanics", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

```{r}

#Most popular names for Males under each Ethnic group

df1 <- top10_names %>% filter(Gender == "MALE", Rank == 1) 

ggplot(df1, aes(x = Year.of.Birth, y = Count, color = Ethnicity, fill = Ethnicity )) +
  geom_line(linetype = "solid", size = 1.5) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  scale_y_continuous(breaks = seq(100, 450, by=50))+
  labs(title = "Most popular names for Males by Ethnicity", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 

#Most popular names for Feales under each Ethnic group

df1 <- top10_names %>% filter(Gender == "FEMALE", Rank == 1) 

ggplot(df1, aes(x = Year.of.Birth, y = Count, fill = Ethnicity, color = Ethnicity )) +
  geom_line(linetype = "solid", size = 1.5) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  scale_y_continuous(breaks = seq(100, 450, by=50))+
  labs(title = "Most popular names for Females by Ethnicity", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

```         
```

```{r}


#A function to find the most popular names of MALES and FEMALE (Not considering the ethnicity)
find_popular_names <- function() {
  
df1 <- top10_names %>% filter(Rank == 1, Gender == "MALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name) %>% 
summarise(freq = sum(Count), .groups = "drop")

df <- data.frame("Child.s.First.Name" = character(0), "freq"  = numeric(0))
years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}
  

df1 <- top10_names %>% filter(Rank == 1, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name) %>% 
  summarise(freq = sum(Count), .groups = "drop")

for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
  }

return(df)
  
}

df <- find_popular_names()






```

```{r}

#Plot to show how the Rank 1 names have varied over the years(Not considering the ethnicity)
ggplot(df, aes(x = Year.of.Birth, y = freq, fill=Gender, color = Gender)) +
  geom_line(linetype = "solid", size = 1.5) +
   scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019))+
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +

  scale_y_continuous(breaks = seq(100, 650, by=50))+
  labs(title = "Most popular names for each Gender", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

```{r}
#Top 5 popular names for Females over the years


find_top5_femaleNames <- function() {
  
df1 <- top10_names %>% filter(Rank == 1, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name, Rank) %>% 
  summarise(freq = sum(Count), .groups = "drop")

df <- data.frame("Child.s.First.Name" = character(0), "freq"  = numeric(0))
years <- unique(df1$Year.of.Birth)

for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}

df1 <- top10_names %>% filter(Rank == 2, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name, Rank) %>% 
  summarise(freq = sum(Count), .groups = "drop")

years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}


df1 <- top10_names %>% filter(Rank == 3, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name, Rank) %>% 
  summarise(freq = sum(Count), .groups = "drop")

years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}


df1 <- top10_names %>% filter(Rank == 4, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name, Rank) %>% 
  summarise(freq = sum(Count), .groups = "drop")

years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}


df1 <- top10_names %>% filter(Rank == 5, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name, Rank) %>% 
  summarise(freq = sum(Count), .groups = "drop")

years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}
  



return(df)
  
}

df <- find_top5_femaleNames()

df$Rank <- as.factor(df$Rank)


```

```{r}
#Plot to show how the Rank 1 names have varied over the years(Not considering the ethnicity)
ggplot(df, aes(x = Year.of.Birth, y = freq, fill=Rank, color = Rank)) +
  geom_line(linetype = "solid", size = 1.5) +
   scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019))+
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3, check.overlap=T) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +

  scale_y_continuous(breaks = seq(100, 650, by=50))+
  labs(title = "Most popular Female names over the years", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

```{r, echo=FALSE}
unique(BabyNames$Ethnicity)
length(unique(BabyNames$Ethnicity))

unique(modified_data$Ethnicity)
length(unique(modified_data$Ethnicity))

unique(babynames$Ethnicity)



```

```{r}
#Relationship between the Rank and the Count variable
ggplot(modified_data, aes(x = Rank, y = Count)) +
  geom_point(color = "lightblue") +
  labs(title = "Rank vs the Count",
       subtitle = "Are names with higher frequency always rank higer?",
       x = "Ranks",
       y = "Count") +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  scale_x_discrete(limits = c(1:102), breaks = seq(1,105, by = 10))


```

```{r}

ggplot(modified_data, aes(x = Rank)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Rank Variable",
       x = "Rank Variable",
       y = "Frequency")
```

```{r}
#Find similar phonetics of name



soundex(modified_data$Child.s.First.Name[1]) == soundex(modified_data$Child.s.First.Name[2]) 

#Add a column with their phonix
modified_data$ph <- soundex(modified_data$Child.s.First.Name)
modified_data$ph1 <- phonex(modified_data$Child.s.First.Name)

phonex(modified_data$Child.s.First.Name[1])
soundex(modified_data$Child.s.First.Name[1])

#Obtain the summary of baby names with simmialar
modified_data %>% filter(Gender=="MALE") %>% 
  group_by(Gender,ph) %>% summarise(count=n()) %>% 
  ggplot(aes(x=count)) +
  geom_histogram()

modified_data %>% filter(Gender=="FEMALE") %>% 
  group_by(Gender,ph) %>% summarise(count=n())

modified_data %>% filter(ph=="A100") %>% select(Year.of.Birth,Ethnicity, Child.s.First.Name)




```

```{r}
# Example: Visualizing distribution of Soundex codes
soundex_freq <- table(modified_data$ph)
barplot(soundex_freq, main = "Distribution of Soundex Codes", xlab = "Soundex Codes", ylab = "Frequency")

soundex_freq <-modified_data %>% filter(Gender=="FEMALE") %>% 
              group_by(Gender,ph) %>% summarise(count=n()) %>% 
              arrange(desc(count))


modified_data %>% filter(Gender=="MALE") %>% 
              group_by(Gender,ph) %>% summarise(count=n()) %>% 
              arrange(desc(count))
modified_data %>% filter(Gender=="FEMALE") %>% 
              group_by(Gender,ph) %>% summarise(count=n()) %>% 
              arrange(desc(count))

d1 <- modified_data %>% filter(ph %in% c("J200","A350"))


#,"K400","L400"


ggplot(soundex_freq, aes(x = ph, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Frequencies of Female", x = "Female", y = "Frequency")




```

```{r}

```

```{r}
phonetic_comparison <- by(modified_data$ph, modified_data$Ethnicity, summary)
print(phonetic_comparison)
```

```{r}
# Example: Visualizing phonetic representations by ethnicity
library(ggplot2)

# Assuming the Phonetic column contains numeric data for visualization
baby_names_data$Phonetic <- as.numeric(modified_data$ph)

ggplot(modified_data, aes(x = Gender, y = ph, color = Gender)) +
  geom_boxplot() +
  labs(title = "Phonetic Representations by Ethnicity", x = "Gender", y = "Phonetic Value")

```
