---
title: "STAT850 - BABY NAMES"
author: "Samitha Herath & Fangfang Zhang"
format: revealjs
scrollable: true
toc: false
---

------------------------------------------------------------------------

## Description of the data set {.smaller}

-   This dataset, `BabyNames`, consist of the popular baby names by sex and ethnic group from 2011 to 2019. It was collected through civil birth registration system.

-   The dataset includes 6 columns:

    -   `Year of Birth`: a quantitative variable indicating the baby's year of birth, ranging from 2011 to 2019
    -   `Gender`: a categorical variable with two categories-Male and Female
    -   `Ethnicity`: a categorical indicating the mother's ethnicity
    -   `Child.s.First.Name`: a categorical variable representing the baby's first name
    -   `Count`: a quantitative variable representing the frequency with which names were collected through civil birth registration
    -   `Rank`: a categorical variable representing the frequency of baby names in descending order for each year.

    The first few observations of the data set

```{r}
BabyNames <- read.csv("Popular_Baby_Names.csv")
head(BabyNames)
```

## Methodology {.smaller}

### Cleaning the data

```{r, echo=FALSE, message=FALSE}
#Required libraries
library(tidyverse)
library(reshape2)
library(phonics)
library(wordcloud)
library(RColorBrewer)

library(ggwordcloud)
```

```{r}

BabyNames <- read.csv("Popular_Baby_Names.csv")



BabyNames$Child.s.First.Name <- tolower(BabyNames$Child.s.First.Name)

babynames<- unique(BabyNames) 

modified_data <- babynames %>% mutate(Ethnicity = recode(Ethnicity, "BLACK NON HISP" = "BLACK NON HISPANIC" ,
                                                      "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                                                      "WHITE NON HISP" = "WHITE NON HISPANIC" )) 
```

-   We identified and removed duplicate rows from the initial dataset.
-   There are some ethic groups have the similar names such as `ASIAN AND PACIFIC ISLANDER` and `ASIAN AND PACI`, `BLACK NON HISPANIC` and `BLACK NON HISP` and `WHITE NON HISPANIC` and `WHITE NON HISP`. They are the same ethic groups but with different names, we combine them before we perform analysis

### Methods

-   In subsetting the dataset for different genders and ethnic groups, we primarily used `dplyr` functions `group_by`, `filter`, `summarise`, `arrange` and `mutate`[@dplyr-package].

-   For instance, to generate a dataset of popular baby names for each gender, we first selected males using the `filter` function. Then we applied `group_by` to create the dataset of all males names, followed by `summarize` to count the frequencies of each name's occurrence.

-   To investigate common results (unisex names) between datasets of different genders within the same ethnic group, we used the `inner_join()` function to combine these datasets and identify any overlapping names.

-   For analyzing the length of names, we used the `str_count` function[@stringr-package] to calculate the number of letters in each name. We also applied `mean` function to calculate the average name lenght for each gender from year 2011 to 2019.

-   For data visualization, `ggplot` function[@ggplot2-package].

-   Creating a word cloud using the `wordcloud` function [@worldcloud-package] providing a graphical representation of the popular names.

## Data Analysis {.smaller}

```{r}
#| echo: false

# select top 10 female's names
vf_data <- modified_data %>% 
  filter(Gender == "FEMALE") %>% 
  group_by(Child.s.First.Name, Gender) %>% 
  summarise(freq = sum (Count),  .groups = "drop") %>% 
  arrange(desc(freq)) %>% head(10)

# select top 10 male's names
vm_data <- modified_data %>% 
  filter(Gender == "MALE") %>% 
  group_by(Child.s.First.Name, Gender) %>%
  summarise(freq = sum(Count), .groups = "drop") %>% 
  arrange(desc(freq)) %>% 
  head(10)
```

```{r}
library(ggplot2)

vm_data$Child.s.First.Name <-factor(vm_data$Child.s.First.Name, levels = vm_data$Child.s.First.Name) 

ggplot(data=vm_data, aes(x= Child.s.First.Name, y=freq)) +
  geom_bar(stat="identity", fill = 'steelblue2') +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(title = "Top 10 Popular Male Babies' Names", 
       subtitle = "from 2011 to 2019", 
       x = "Child's First Name", 
       y = "Count") + 
    theme_minimal()

vf_data$Child.s.First.Name <-factor(vf_data$Child.s.First.Name, levels = vf_data$Child.s.First.Name) 

ggplot(data=vf_data, aes(x= Child.s.First.Name, y=freq)) +
  geom_bar(stat="identity", fill = 'hotpink') +
  theme(axis.text.x = element_text(angle =45, hjust = 1)) +
  labs(title = "Top 10 Popular Female Babies' Names", 
       subtitle = "from 2011 to 2019", 
       x = "Child's First Name", 
       y = "Count") + 
    theme_minimal()
```


-   We can see the top 10 names for each gender on horizontal axis from these two bar charts, and the frequency of each name displayed on vertical axis.



Display the proportions of the name ‘Olivia’ in both Black non-Hispanic and White non-Hispanic groups

```{r calculate ratio of Olivia in each ethnic group}
#| echo: false
#| 

total_ef2 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% summarise(total_BNH = sum(Count))

total_count <- total_ef2$total_BNH

freq_ef2 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_BNH = sum(Count)) %>% ungroup()

prop_ef2 <- freq_ef2 %>% mutate(prop_F_BNH = freqF_BNH / total_count) %>% filter(Child.s.First.Name == 'olivia')
##-------------------------------------------------------------###

total_ef4 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% summarise(total_WNH = sum(Count))

total_count4 <- total_ef4$total_WNH

freq_ef4 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "FEMALE") %>% group_by(Child.s.First.Name) %>% summarise(freqF_WNH = sum(Count)) %>% ungroup()

prop_ef4 <- freq_ef4 %>% mutate(prop_F_WNH = freqF_WNH / total_count4) %>% filter(Child.s.First.Name == 'olivia')

propData_ef2 <- data.frame(prop_ef2$Child.s.First.Name, prop_ef2$prop_F_BNH , prop_ef4$prop_F_WNH)
colnames(propData_ef2) <- c("Child_Name", "Black non-Hispanic - Female", "White non-Hispanic - Female")

```


```{r}
melt_prop2 <- melt(propData_ef2, id.vars = 'Child_Name')
melt_prop2$variable <- as.character(melt_prop2$variable )


ggplot(data = melt_prop2, aes(x = Child_Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Proportion of Olivia Between Black non-Hispanic and White non-Hispanic",
       x = "Name",
       y = "Proportion",
       fill = "Ethnic Group")+
  theme_minimal()
```

-  Interestingly, it appears that the White non-Hispanic group has a stronger preference for this name.


```{r calculate ratio of commonlarity in WNH vs BNH for males}
#| echo: false
#| 



total_em2 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% summarise(total_BNH = sum(Count))

total_count_em2 <- total_em2$total_BNH

freq_em2 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "BLACK NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_BNH = sum(Count)) %>% ungroup()

prop_em2 <- freq_em2 %>% mutate(prop_M_BNH = freqM_BNH / total_count) %>% 
  select(Child.s.First.Name, prop_M_BNH) %>%  arrange(desc(prop_M_BNH))


##-------------------------------------------------------------###

total_em4 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% summarise(total_WNH = sum(Count))

total_count_em4 <- total_em4$total_WNH

freq_em4 <- modified_data %>% select('Ethnicity', 'Gender' , 'Child.s.First.Name', 'Count') %>% filter(Ethnicity == "WHITE NON HISPANIC", Gender == "MALE") %>% group_by(Child.s.First.Name) %>% summarise(freqM_WNH = sum(Count)) %>% ungroup()

prop_em4 <- freq_em4 %>% mutate(prop_M_WNH = freqM_WNH / total_count4) %>% 
  select(Child.s.First.Name, prop_M_WNH) %>% arrange(desc(prop_M_WNH))



##-------------------------------------------------------------###

inner_em4 <- inner_join(prop_em2, prop_em4, by = "Child.s.First.Name") %>% 
  filter(Child.s.First.Name %in%  c('david', 'jacob', 'michael', 'noah'))

colnames(inner_em4) <- c("Child_Name", "Black non-Hispanic - Male", "White non-Hispanic - Male")

```




```{r}
melt_prop4 <- melt(inner_em4, id.vars = 'Child_Name')
melt_prop4$variable <- as.character(melt_prop4$variable )

ggplot(data = melt_prop4, aes(x = Child_Name, y = value, fill = variable)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  labs(title = "Proportion of Males Names Between Black non-Hispanic and White non-Hispanic",
       subtitle = "within top 20", 
       x = "Name",
       y = "Proportion",
       fill = "Ethnic Group")+
  theme_minimal()
```

-  BNH prefers 'Noah' over WNH for naming a male baby. 



The following figures display the average length names for each gender and ethnic group.

```{r calculate average length of letters for names}
#| echo: false

library(stringr)
# calculate the length of letter of each name
length_data <- modified_data %>% mutate(N.Length = str_count(Child.s.First.Name, "[A-Za-z]")) %>% mutate(T.length = Count * N.Length) 
# calculate the average number of letters in each year for males and females
avglength_data <- length_data  %>% group_by(Year.of.Birth, Gender) %>% 
  summarise(Avg.Letters = mean(N.Length, na.rm = TRUE), .groups = "drop") %>% 
  arrange(Year.of.Birth, Gender)
# calculate the average number of letters in each year for each ethnic groups
avglength_data_ethnic <- length_data  %>% group_by(Year.of.Birth, Ethnicity) %>% 
  summarise(Avg.Letters = mean(N.Length, na.rm = TRUE), .groups = "drop") %>% 
  arrange(Year.of.Birth, Ethnicity)

```



```{r}
#| label: fig-AverageLength1
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "Average Length of First Names for Males and Females"
#| echo: false


ggplot(data = avglength_data, aes(x = Year.of.Birth, y = Avg.Letters, color = Gender, group = Gender)) + 
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2011, 2019, by=1), limits = c(2011, 2019)) +
  labs(title = "Average Number of Letters in Babies' Name by Gender", 
       subtitle = "from 2011 to 2019", 
       x = "Year", 
       y = "Average Number of Letters") + 
    theme_minimal()
```


```{r}
#| label: fig-AverageLength2
#| fig-width: 8
#| fig-height: 4 
#| out-width: 80%  
#| fig-cap: "Average Length of First Names for Different Ethnic Groups"
#| echo: false


ggplot(data = avglength_data_ethnic, aes(x = Year.of.Birth, y = Avg.Letters, color = Ethnicity, group = Ethnicity)) + 
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(2011, 2019, by=1), limits = c(2011, 2019)) +
  labs(title = "Average Number of Letters in Babies' Name by Ethnicity", 
       subtitle = "from 2011 to 2019", 
       x = "Year", 
       y = "Average Number of Letters") + 
    theme_minimal()
```


-   overall, the average length of names has generally decreased over time 
-   trend is a growing preference for shorter names 
-   "Short Names Are Long on Earning Power"


## Data Analysis Continued {.smaller}

-   The following plots clearly show the trend of the most popular baby names for both males and females over the years for each ethnic group.

```{r}
top10_names <- modified_data %>% filter(Rank %in% c(1:10))
```

```{r}
df1 <- top10_names %>% filter(Gender == "MALE", Rank == 1) 

ggplot(df1, aes(x = Year.of.Birth, y = Count, color = Ethnicity, fill = Ethnicity )) +
  geom_line(linetype = "solid", size = 1.5) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  scale_y_continuous(breaks = seq(100, 450, by=50))+
  labs(title = "Most popular names for males by Ethnicity", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

```{r}
df1 <- top10_names %>% filter(Gender == "FEMALE", Rank == 1) 

ggplot(df1, aes(x = Year.of.Birth, y = Count, fill = Ethnicity, color = Ethnicity )) +
  geom_line(linetype = "solid", size = 1.5) +
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +
  scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019)) +
  scale_y_continuous(breaks = seq(100, 450, by=50))+
  labs(title = "Most popular names for females by Ethnicity", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency")
```

*Through these plots, we visualize the dynamic trends in name popularity, revealing notable spikes or dips in specific years. They showcase the evolving patterns in name selection, providing a comprehensive view of how naming preferences evolve over time.*

-   The next plot illustrates the most popular names for babies, regardless of their gender, across different years.

```{r}
#A function to find the most popular names of MALES and FEMALE (Not considering the ethnicity) 

#Note: Planning on trying to improve this function.
find_popular_names <- function() {
  
df1 <- top10_names %>% filter(Rank == 1, Gender == "MALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name) %>% 
summarise(freq = sum(Count), .groups = "drop")

df <- data.frame("Child.s.First.Name" = character(0), "freq"  = numeric(0))
years <- unique(df1$Year.of.Birth)
  
for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
}
  

df1 <- top10_names %>% filter(Rank == 1, Gender == "FEMALE" ) %>% 
group_by(Year.of.Birth,Gender,Child.s.First.Name) %>% 
  summarise(freq = sum(Count), .groups = "drop")

for (i in 1:length(years)) {
  df0 <- df1 %>% filter(Year.of.Birth == years[i])
    
  row <- df0[df0$freq == max(df0$freq), ]
    
  df <- rbind(df, row)
    
  }

return(df)
  
}

df <- find_popular_names()






```

```{r, warning=FALSE}
ggplot(df, aes(x = Year.of.Birth, y = freq, fill=Gender, color = Gender)) +
  geom_line(linetype = "solid", size = 1.5) +
   scale_x_discrete(limits = c(2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018,2019))+
  geom_label(aes(label = c(Child.s.First.Name)), nudge_x = 0.25, nudge_y = 0.15, color = "blue", size = 3) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +

  scale_y_continuous(breaks = seq(100, 650, by=50))+
  labs(title = "Most popular names for each Gender", subtitle = "(2011 - 2019)", x = "Birth Year", y = "Frequency") 
```

-   *The names **Jayden** and **Liam** held the top positions as the most popular choices for baby boys across diverse ethnicities from 2011 to 2019. Meanwhile, **Isabella**, **Olivia** , and **Esther** emerged as the favored names for baby girls during the same period. This trend persisted regardless of cultural backgrounds.*

-   *A search engine query for the most popular baby names in the United States consistently showcased **Liam** as the leading choice for boys and **Olivia** as the foremost pick for girls. Our analysis aligns with these findings, reflecting a parallel preference for these names across the nation.*

**Word cloud**

Creating a word cloud for the popular baby names for each gender

```{r word cloud}


modified_data %>% 
  filter(Gender == "MALE" & Rank %in% c(1:100)) %>% 
  group_by(Child.s.First.Name) %>% 
  summarise(freq = sum(Count)) %>% arrange(desc(freq)) %>% 
  head(100) %>% 
  with(wordcloud(Child.s.First.Name, freq, max.words = 100, 
                 scale = c(2, 0.5), 
                 colors = brewer.pal(8, "Dark2")))
title(main = "Most popolar baby names for males")
```

```{r}



modified_data %>% 
  filter(Gender == "FEMALE" & Rank %in% c(1:100)) %>% 
  group_by(Child.s.First.Name) %>% 
  summarise(freq = sum(Count)) %>% arrange(desc(freq)) %>% 
  head(100) %>% 
  with(wordcloud(Child.s.First.Name, freq, max.words = 100, 
                 scale = c(2, 0.5), 
                 colors = brewer.pal(8, "Dark2")))
title(main = "Most popolar baby names for females")
```

-   *A word cloud is a visual representation of text data. Typically, the more frequently a word appears in the source material, the larger and more prominently it is displayed in the word cloud.*

-   *The word clouds displaying popular names for boys and girls highlight a clear trend: a small group of names appears frequently, while many others show up less often. This suggests that a handful of names are super popular, while the rest are not as commonly used in the dataset.*

## Conclusion {.smaller}

-   Different ethnic groups, particularly White Non-Hispanic and Black Non-Hispanic, use different baby names, and it reflects diverse cultures. 

-   There has been a noticeable decrease in the average length of names over time, with a growing preference for shorter names. 

-   Across diverse ethnicities from 2011 to 2019, the names `Jayden` and `Liam` were widely favored for baby boys, while `Isabella`, `Olivia`, and `Esther` were popular choices for baby girls. 

-   A consistent trend observed through search engine queries on popular baby names in the United States reaffirmed Liam's prominence among boys and Olivia's popularity among girls. 
